
%\documentclass[mathserif]{beamer}
\documentclass[handout]{beamer}
%\usetheme{Goettingen}
%\usetheme{Warsaw}
\usetheme{Singapore}



%\usetheme{Frankfurt}
%\usetheme{Copenhagen}
%\usetheme{Szeged}
%\usetheme{Montpellier}
%\usetheme{CambridgeUS}
%\usecolortheme{}
%\setbeamercovered{transparent}
\usepackage[english, activeacute]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb}
\usepackage{dsfont}
\usepackage{graphics}
\usepackage{cases}
\usepackage{graphicx}
\usepackage{pgf}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{multirow}	
\usepackage{amstext}
\usepackage[ruled,vlined,lined]{algorithm2e}
\usepackage{amsmath}
\usepackage{epic}
\usepackage{epsfig}
\usepackage{fontenc}
\usepackage{framed,color}
\usepackage{palatino, url, multicol}
%\algsetup{indent=2em}
\newcommand{\factorial}{\ensuremath{\mbox{\sc Factorial}}}
\newcommand{\BIGOP}[1]{\mathop{\mathchoice%
{\raise-0.22em\hbox{\huge $#1$}}%
{\raise-0.05em\hbox{\Large $#1$}}{\hbox{\large $#1$}}{#1}}}
\newcommand{\bigtimes}{\BIGOP{\times}}
\vspace{-0.5cm}
\title{Natural Language Processing \\ MEMMs and CRFs}
\vspace{-0.5cm}
\author[Felipe Bravo Márquez]{\footnotesize
%\author{\footnotesize  
 \textcolor[rgb]{0.00,0.00,1.00}{Felipe Bravo-Marquez}} 
  
 

\date{\today}

\begin{document}
\begin{frame}
\titlepage


\end{frame}


\begin{frame}{MEMMs}
\begin{scriptsize}
\begin{itemize}

\item The goal of sequence labeling is to assign tags to words, or more generally, to assign discrete labels to discrete elements in a sequence \cite{jacobbook}.

\item Well known examples of this problem are: part-of-speech tagging (POS) and Named Entity Recognition (NER).

\item Maximum-entropy Markov models (MEMMs) make use of log-linear models for sequence labeling tasks.
 
 \item In the early NLP literature, logistic regression was often called maximum entropy classification \cite{jacobbook}.
 
 \item Hence, MEMMs will look very similar to the multi-class softmax models seen in the lecture about linear models. 
 
  \item In contrast to HMMs, here we rely on parameterized functions.

\end{itemize}


\end{scriptsize}
\end{frame}



\begin{frame}{MEMMs}
\begin{scriptsize}
\begin{itemize}


 \item The goal of MEMMs is  model the following conditional distribution:
 
 \begin{displaymath}
  P(s_1,s_2 \dots, s_m | x_1, \dots, x_m)
 \end{displaymath}

 \item Where each $x_j$ for $j = 1 \dots m$ is the $j$-th input symbol (for example the j-th word in a sentence), and each $s_j$ for $j = 1 \dots m$ is the $j$-th tag.\footnote{These slides are based on lecture notes of Michael Collins \url{http://www.cs.columbia.edu/~mcollins/}. The notation and terminology has been adapted to be consistent with the other material.}

\item We expecto this $P$(DET,NOUN,VERB$|$the,dog,barks$)$ to be higher than $P$(VERB,VERB,VERB$|$the,dog,barks$)$
 
\end{itemize}


\end{scriptsize}
\end{frame}


\begin{frame}{MEMMs}
\begin{scriptsize}
\begin{itemize}
\item We use $S$ to denote the set of possible tags.
\item We assume that $S$ is a finite set. 
 \item For example, in part-of-speech tagging of English, $S$ would be the set of all possible parts of speech in English (noun, verb, determiner, preposition, etc.).
 \item Given a sequence of words $x_1, \dots, x_m$ , there are $k^m$ possible part-of-speech sequences $s_1, \dots, s_m$ , where $k = |S|$ is the number of possible parts of speech.
 \item We want to estimate a distribution over these $k^m$ possible sequences.
\end{itemize}


\end{scriptsize}
\end{frame}


\begin{frame}{MEMMs}
\begin{scriptsize}
\begin{itemize}
 \item In a first step, MEMMs use the following decomposition:
 \begin{equation}
\begin{split}
  P(s_1,s_2 \dots, s_m | x_1, \dots, x_m) \quad & =  \prod_{i=1}^{m}    P(s_i | s_1 \dots, s_{i-1}, x_1, \dots, x_m)\\
 \quad & =  \prod_{i=1}^{m}    P(s_i | s_{i-1}, x_1, \dots, x_m)
\end{split}
\end{equation}

\item The first equality is exact (it follows by the chain rule of conditional probabilities).

\item The second equality follows from an independence assumption, namely that for all $i$,

\begin{displaymath}
 P(s_i | s_1 \dots, s_{i-1}, x_1, \dots, x_m) =   P(s_i | s_{i-1}, x_1, \dots, x_m)
\end{displaymath}


 
\end{itemize}


\end{scriptsize}
\end{frame}


\begin{frame}{MEMMs}
\begin{scriptsize}
\begin{itemize}
 \item Hence we are making a first order Markov assumption similar to the Markov assumption made in HMMs. 
 
 \item The tag in the $i$-th position depends only on the tag in the $(i -1)$-th position. 
 
 
 \item Having made these independence assumptions, we then model each term using a log-linear (Softmax) model:
 
 \begin{equation}
 P(s_i | s_{i-1}, x_1, \dots, x_m)  =  \frac{\exp (\vec{w}\cdot \vec{\phi}(x_1, \cdots, x_m, i, s_{i-1},s_i))}{\sum_{s' \in S} \exp (\vec{w}\cdot \vec{\phi}(x_1, \cdots, x_m, i, s_{i-1},s'))}
\end{equation}


 
\end{itemize}


\end{scriptsize}
\end{frame}



\begin{frame}{MEMMs}
\begin{scriptsize}

Here $\vec{\phi}(x_1, \cdots, x_m, i, s_{i-1},s_i)$ is a feature vector where:
\begin{itemize}
 \item $x_1, \cdots, x_m$ is the entire sentence being tagged.
  \item $i$ is the position to be tagged (can take any value from $1$ to $m$)
  \item $s$ is the previous tag value (can take any value in $S$).
  \item $s'$ is the new tag value (can take any value in $S$)
 
\end{itemize}


\end{scriptsize}
\end{frame}

\begin{frame}{Example of Features used in Part-of-Speech Tagging}
\begin{scriptsize}


\begin{enumerate}
 \item $\vec{\phi}(x_1, \cdots, x_m, i, s_{i-1},s_i)_{[1]}=1$ if $s_i$ = ADVERB and word $x_i$ ends in ``-ly''; 0 otherwise. \\ 
 
 If the weight $\vec{w}_{[1]}$ associated with this feature is large and positive, then this feature is essentially saying that we prefer labelings where words ending in -ly get labeled as ADVERB.
 
 \item $\vec{\phi}(x_1, \cdots, x_m, i, s_{i-1},s_i)_{[2]}=1$ if $i=1$, $s_i$= VERB, and $x_m$=?; 0 otherwise. 
 \\ If the weight $\vec{w}_{[2]}$ associated with this feature is large and positive, then labelings that assign VERB to the first word in a question (e.g., ``Is this a sentence beginning with a verb?'') are preferred.


\item $\vec{\phi}(x_1, \cdots, x_m, i, s_{i-1},s_i)_{[3]}=1$ if $s_{i-1}$= ADJECTIVE and $l_i$= NOUN; 0 otherwise. 
\\Again, a positive weight for this feature means that adjectives tend to be followed by nouns. 

\item $\vec{\phi}(x_1, \cdots, x_m, i, s_{i-1},s_i)_{[4]}=1$ if $s_{i-1}$= PREPOSITION and $s_{i}$= PREPOSITION. 
\\ A negative weight $\vec{w}_{[4]}$ for this function would mean that prepositions don't tend to follow prepositions.

 
\end{enumerate}


\footnotemark{Source: \url{https://blog.echen.me/2012/01/03/introduction-to-conditional-random-fields/}}

\end{scriptsize}
\end{frame}



\begin{frame}{Feature Templates}
\begin{scriptsize}

It is possible to define more general feature templates covering unigrams, bigrams, n-grams of words as well as tag values of $s_{i-1}$ and $s_i$.

\begin{enumerate}
  
 \item A word unigram and tag unigram feature template: $\vec{\phi}(x_1, \cdots, x_m, i, s_{i-1},s_i)_{[hash(j,k)]}=1$ if $s_i$ = TAG$_{[j]}$ and $x_i$ = WORD$_{[k]}$; 0 otherwise $\forall j,k$. \\ Notice that $j$ is and index spanning all possible tags in $S$ and $k$ is another index spanning the words in the vocabulary $V$.
 
 \item A word bigram and tag bigram feature template: $\vec{\phi}(x_1, \cdots, x_m, i, s_{i-1},s_i)_{[hash(j,k,u,v)]}=1$ if $s_{i-1}$ = TAG$_{[j]}$ and $s_i$ = TAG$_{[k]}$ and $x_{i-1}$ = WORD$_{[u]}$ and $x_{i}$ = WORD$_{[v]}$; 0 otherwise $\forall j,k,u,v$. 
 
 
\end{enumerate}

The function $hash(j,k,...)$ will map each different feature to a unique index in the feature vector.  \\
Notice that the resuling vector will be very high-dimensional and sparse.
\end{scriptsize}
\end{frame}



\begin{frame}{MEMMs and Multi-class Softmax}
\begin{scriptsize}
\begin{itemize}
\item Notice that the log-linear model from above is very similar to the multi-class softmax model presented in the lecture about linear models.

\item A general log-linear model has the following form:

\begin{displaymath}
 P( y | x; \vec{w}) = \frac{\exp (\vec{w}\cdot \vec{\phi}(x,y))}{\sum_{y' \in Y} \exp (\vec{w}\cdot \vec{\phi}(x,y'))}
\end{displaymath}


\item A multi-class softmax model has the following form:
\begin{equation}
\begin{split}
\hat{\vec{y}} \quad & =  \operatorname{softmax}(\vec{x} \cdot W + \vec{b})  \\
\hat{\vec{y}}_{[i]} \quad & = \frac{e^{(\vec{x} \cdot W + \vec{b})_{[i]}}}{\sum_j e^{(\vec{x} \cdot W + \vec{b})_{[j]}}}
\end{split}
\end{equation}

 
\end{itemize}


\end{scriptsize}
\end{frame}




\begin{frame}{MEMMs and Multi-class Softmax}
\begin{scriptsize}
\begin{itemize}

\item Difference 1: in the log-linear model we have a fixed parameter vector $\vec{w}$ instead of having multiple vectors (one column of $W$ for each class value).

\item Difference 2: the feature vector of the log-linear model $\vec{\phi}(x,y)$ includes information of the label $y$, whereas the input vecotr $\vec{x}$ of the softmax model is independent of $y$. 

\item Log-linear models allow using features that consider the interaction between $x$ and $y$ (e.g., $x$ end in ``ly'' and $y$ is an ADVERB).

 
\end{itemize}


\end{scriptsize}
\end{frame}



\begin{frame}{Training MEMMs}
\begin{scriptsize}
\begin{itemize}

\item Once we've defined the feature vectors $\vec{\phi}$, we can train the parameters $\vec{w}$ of the model in the usual way for linear models.

\item We set the negative log-likelihood as the loss function and optimize parameters using gradient descent from the training examples.

\item This is equivalent as using the cross-entropy loss.

\item ``Any loss consisting of a negative log-likelihood is a cross-entropy between the empirical distribution defined by the training set and the probability distribution defined by model'' \cite{goodfellow2016deep}.  
 
\end{itemize}


\end{scriptsize}
\end{frame}




\begin{frame}{Decoding with MEMMs}
\begin{scriptsize}
\begin{itemize}

\item The decoding problem is as follows.
\item We are given a new test sequence $x_1, \dots, x_m$.
\item Our goal is to compute the most likely state sequence for this test sequence,

\begin{equation}
 \operatorname{argmax}_{s_1,\dots,s_m} P(s_1,\dots,s_m|x_1,\dots,x_m).
\end{equation}

\item There are $k^m$ possible state sequences, so for any reasonably large sentence length $m$ brute-force search through all the possibilities will not be possible.

\item We can use the Viterbi alogrithm in a similar way as used for HMMs.
 
\end{itemize}


\end{scriptsize}
\end{frame}



\begin{frame}{Decoding with MEMMs}
\begin{scriptsize}
\begin{itemize}

\item The basic data structure in the algorithm will be a dynamic programming table $\pi$ with entries $\pi[j,s]$ for $j=1, \dots, m$, and $s \in S$.

\item  $\pi[j,s]$ will store the maximum probability for any state sequence ending in state $s$ at position $j$.

\item More formally, our algorithm will compute 
\begin{displaymath}
\pi[j,s] =  \operatorname{max}_{s_1,\dots, s_{j-1}}\left(P(s | s_{j-1}, x_1, \dots, x_m) \prod_{k=1}^{j-1}    P(s_k | s_{k-1}, x_1, \dots, x_m)\right)
\end{displaymath}
for all $j = 1, \dots,m$, and for all $s \in S$.

\end{itemize}


\end{scriptsize}
\end{frame}



\begin{frame}{Links}
\begin{scriptsize}
\begin{itemize}

\item \url{https://www.depends-on-the-definition.com/sequence-tagging-lstm-crf/}

\item \url{https://www.quora.com/What-are-the-pros-and-cons-of-these-three-sequence-models-MaxEnt-Markov-Model-Conditional-random-fields-and-recurrent-neural-networks}

\item \url{https://people.cs.umass.edu/~mccallum/papers/crf-tutorial.pdf}

\end{itemize}


\end{scriptsize}
\end{frame}






\begin{frame}
\frametitle{Questions?}
%\vspace{1.5cm}
\begin{center}\LARGE Thanks for your Attention!\\ \end{center}



\end{frame}

\begin{frame}[allowframebreaks]\scriptsize
\frametitle{References}
\bibliography{bio}
\bibliographystyle{apalike}
%\bibliographystyle{flexbib}
\end{frame}  


%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
